Avoiding tipping points in fisheries management: A Bayesian Non-Parametric approach

### 2014-06-23 manuscripts/manuscript.Rmd
additional citation added to references.bib, keep_tex = true
marc's edits integrated

abstract: |
          Model uncertainty and limited data coverage are fundamental
          challenges to robust management of human intervention in a 
          natural system.  These challenges
          are acutely highlighted by concerns that many ecological systems
          may contain tipping points, such as Allee population sizes.  
          Before a collapse, we do not know
          where the tipping points lie, if they exist at all.  Hence,
          we know neither a complete model of the system dynamics nor
          do we have access to data in some large region of state-space
          where such a tipping point might exist.  We illustrate
          how a Bayesian Non-Parametric (BNP) approach using a Gaussian Process (GP) prior
          provides a flexible representation of this inherent
          uncertainty. We embed GPs in a
          stochastic dynamic programming framework in order to make
          robust management predictions with both model uncertainty and
          limited data. We conduct a set of simulations to evaluate this
          approach as compared with the standard approach of using model
          selection to choose from a set of candidate models.  We find that model selection erroneously
          favors models without tipping points -- leading to harvest policies
          that guarantee extinction. The GP approach
          performs nearly as well as the true model and significantly
          outperforms standard appraoches. We illustrate this using examples
          of simulated single-species dynamics, where the standard model
          selection approach should be most effective, and find that it
          still fails to account for uncertainty appropriately and leads
          to population crashes, while management based on the GP 
          approach does not, since it does not underestimate the uncertainty
          outside of the observed data.

@@

Decision making under uncertainty is a ubiquitous challenge in the
management of human intervention in natural resources and conservation.
Ecological dynamics are frequently complex and difficult to measure,
making uncertainty in our understanding a persistent challenge.
Decision-theoretic approaches provide a framework to determine the
best sequence of actions in face of uncertainty, but only when that
uncertainty can be meaningfully quantified [@Fischer2009].  Over the last
four decades (but stretching even further back in time), marked by  the
seminal contributions of @Clark1976; @Clark2009 and @Walters1978,
dynamic optimization methods, particularly Stochastic Dynamic Programming
(SDP, @Clark2009), have become increasingly important as a means of
understanding how to management human intervention into natural systems.
Concomittantly, there has been increasing recognition of the importance of
Allee population sizes or `tipping points' [@Scheffer2001;
@Polasky2011] that management actions may inadvertently
cause populations to cross.  @Scheffer2009 emphasized the difficulties of
formulating even qualitatively correct models of the underlying processes.

We develop a novel approach to address these concerns in the context of
fisheries; though the underlying challenges and methods are germane to
many other problems of conservation or natural resource exploitation,
such as forestry.  Economic value and ecological concern have made
marine fisheries the crucible for much of the founding work [@Gordon1954;
@Clark1976; -@Clark2009; @Reed1979; @May1979; @Ludwig1982] on management
under uncertainty.  Global declines [@Worm2006] and the controversy
surrounding their interpretation [@Hilborn2007; @Worm2009] have made
understanding these challenges all the more pressing.

@@

The dual concerns of model uncertainty and
incomplete data coverage pose a substantial challenge to existing decision-theoretic
approaches [@Brozovic2011].  Because intervention is often (but not always, see @Hughes2013) too late after a tipping point has been crossed, management
is most often concerned with avoiding potentially catastrophic tipping
points before any data is available at or following a transition that
would more clearly reveal these regime shift dynamics [e.g. @Bestelmeyer2012].

We illustrate how a stochastic dynamic programming (SDP) algorithm
[@Mangel1988; @Marescot2013] can be driven by the
predictions from a Bayesian Non-Parametric (BNP) model of population dynamics [@Munch2005a].
This provides two distinct advantages compared with contemporary
approaches.  First, using a BNP sidesteps the need for an accurate
model-based description of the system dynamics.  Second, the BNP can
better reflect uncertainty that arises when extrapolating a model outside
of the data on which it was fit.  We illustrate that when the correct
model is not known, this latter feature is crucial to providing a robust
decision-theoretic approach in face of substantial structural uncertainty.

This paper represents the first ecological application of the SDP
decision-making framework without an a priori model of the underlying
dynamics.  In contrast to parametric models which can only reflect
uncertainty in parameter estimates, the BNP approach provides a more
state-space dependent representation of uncertainty.  This permits
a much greater uncertainty far from the observed data than near the
observed data.  These features allow the Gaussian Process Dynamic Programming (GPDP) approach to find robust
management solutions in face of limited data and without knowledge of
the correct model structure.

@@

We let $X_t$ denote the size (numbers or biomass) of the focal population at time $t$ and assume that in the absence of take its dynamics are

$$ X_{t+1}= Z_t f(X_t, h_t, p) $$

@@

In order to find that policy, we introduce the value function$V_t(x_t)$ representing the total discounted catch from time $t$ onwards given that $X_t=x$.  This value function satisfies an equation of Stochastic Dynamic Programming  [SDP, @Mangel1988; @Clark2000; @Clark2009], sometimes called the Bellman-Hamilton-Jacobi equation [@Mangel2014],

@@

where expectation is taken over all possible values of the next state, $X_t{+1}$, and maximized over all possible choices of harvest, $h_t$.  That is, at time $t$, when population size is $x$ and harvest $h_t$ is applied, the immediate return is $R(h_t, X_t)$. In the classical case, the sole source of uncertainty is the process noise term, $Z$, and thus the expectation above is equivalent to taking expectations over $Z$. That is

@@

We consider the finite time problem with $T=$`r MaxT`, which we solve using value iteration algorithm [@Mangel1988; @Clark2000].

@@

The core difference for our purpose between the GP representation and the parametric
models we have already introduced is the "non-parametric" property, which
means that an estimated GP model is defined explicitly in reference to 
some observed data.  Unlike the models above, it cannot be specified by
the value of some parameters alone.  The nonparametric moniker for the
GP approach is unfortunate as (a) it still involves the estimation of
parameters, and (b), statisticians also use non-parametric to mean models
that make only weak distributional assumptions  [e.g. @Lehmann1975].
However, we retain the terminology for consistency with previous studies.

@@

Assuming the data $X_o$ are observed with some process noise,

@@

$\varepsilon$ IID normal, variance $\sigma_n^2$. Note that we have chosen to
assume additive noise.  While we could just as easily consider log-normal noise
as in the parametric models, we make this choice to bring home the point that the
Gaussian process approach need not have structurally correct noise form either.

@@

We use vague inverse Gamma priors on both the lengthscale $\ell$ and process noise $\sigma$ parameters, of the form

@@

Using the simulated training data we also estimate a Gaussian process
defined by a radial basis function kernel of two parameters: $\ell$,
which gives the characteristic length-scale over which correlation between
two points in state-space decays, and $\sigma$, which gives the scale
of the process noise by which observations $Y_{t+1}$ may differ from
their predicted values $X_{t+1}$ given an observation of the previous
state, $X_t$. @Munch2005a gives an accessible introduction to the use
of Gaussian processes in providing a Bayesian nonparametric description
of the stock-recruitment relationship.

@@

We use a Metropolis-Hastings Markov Chain Monte Carlo (@Gelman2003) to
infer posterior distributions of the two parameters of the GP (Figure
S13, code in appendix), under weakly informative Gaussian priors
(see parameters in table S5). As the posterior distributions differ
substantially from the priors (Figure S13), we can be assured that most
of the information in the posterior comes from the data rather than the
prior belief.

@@

The method of Gaussian Process Dynamic Programming
--------------------------------------------------

@@

To ensure our results are robust to our choice of parameters, 
we perform our analysis over 96 different scenarios. To better
understand the process, we will first describe in detail the
results of a single scenario.

@@

All models fit the observed data rather closely and with relatively small uncertainty, as illustrated in the posterior predictive curves in Figure 1.  Figure 1 shows the training data of stock sizes observed over time as points, overlaid with the step-ahead predictions of each estimated model using the parameters sampled from their posterior distributions.  Each model manages to fit the observed data rather closely. Compared to the true model most estimates appear to over-fit, predicting fluctuations that are actually due purely to stochasticity.  All model selection criteria shown in Table 1 penalize more complex models and show a preference for the simpler Ricker model over the more complicated alternate stable state models (Allen and Myers).  Details on MCMC estimates for each model, traces, and posterior distributions can be found in the appendix.

@@

```{r figure_2, dependson=c("assemble-models", "plot-options"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.  Graph shows the expected value for the function $f$ under each model.  Two standard deviations from the estimated Gaussian process covariance with (light grey) and without (darker grey) measurement error are also shown.  The training data is also shown as black points.  The GP is conditioned on (0,0), shown as a pseudo-data point.", fig.width=8, fig.height=6}
x_grid_short <- x_grid[1:40]
gp_short <- gp_predict(gp, x_grid_short, burnin=1e4, thin=300)
models_posteriors <- 
  melt(lapply(sample(100, 50), 
              function(i){
    sample_gp <- mvrnorm(1, 
                            gp_short$Ef_posterior[,i],         
                            gp_short$Cf_posterior[[i]])
    data.frame(stock = x_grid_short, 
               GP = sample_gp,
               y = sample_gp,
               ymin = sample_gp - 2 * sqrt(gp_short$E_Vf), 
               ymax = sample_gp + 2 * sqrt(gp_short$E_Vf), 
               True = sapply(x_grid_short,f,0, p),  
               MLE = sapply(x_grid_short,f,0, est$p), 
               Allen = sapply(x_grid_short, allen_f, 0, pardist[i,]), 
               Ricker = sapply(x_grid_short, ricker_f, 0, ricker_pardist[i,]), 
               Myers = sapply(x_grid_short, myers_f, 0, myers_pardist[i,]))
             }), 
       id=c("stock", "y", "ymin", "ymax"))
ggplot(models_posteriors) + 
    geom_ribbon(aes(x=stock, y=y, ymin=ymin, ymax=ymax, group=L1), 
                  fill = "gray80", 
                  data=subset(models_posteriors, variable == "GP")) + 
    geom_line(aes(stock, value, col = variable, 
                  group=interaction(L1,variable)), 
              alpha=.2) + 
    geom_point(data = obs, aes(x,y), alpha = 0.8) + 
    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
    facet_wrap(~variable) + 
    scale_colour_manual(values=colorkey) +  
    theme(legend.position="none")

@@

The mean inferred state space dynamics of each model relative to the
true model used to generate the data is shown in Figure 2, predicting
the relationship between observed stock size (x-axis) to the stock
size after recruitment the following year.  Note that in contrast to
the  other models shown, the expected Gaussian process corresponds to a
distribution of curves - as indicated by the gray band - which itself
has a mean shown in black. Parameter uncertainty (not shown) spreads
out the estimates further.  The observed data from which each model is
estimated is also shown.  The observations come from only a limited region
of state space corresponding to unharvested or weakly harvested system.
No observations occur at the theoretical optimum harvest rate or near
the tipping point.

@@

Despite the similarities in model fits to the observed data, the
policies inferred under each model differ widely, as shown in Figure 4.
Policies are shown in terms of target escapement, $S_t$.  Under models
such as this a constant escapement policy is expected to be optimal
[@Reed1979], whereby population levels below a certain size $S$ are
unharvested, while above that size the harvest strategy aims to return the
population to $S$, resulting in the hockey-stick shaped policies shown.
Only the structurally correct model (Allen model) and the GP produce
policies close to the true optimum policy (where both the underlying
model structure and parameter values are known without error).

```{r figure_5, dependson=c("sim"), fig.cap="In the management context, GPDP outperforms approaches based on parametric models. Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6}
ggplot(sims_data) + 
  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) + 
  facet_wrap(~method) + guides(legend.position="none")

@@

The consequences of managing 100 replicate realizations of the simulated
fishery under each of the policies estimated is shown in Figure 5.
As expected from the policy curves, the structurally correct model
under-harvests, leaving the stock to vary around it's un-fished optimum.
The structurally incorrect Ricker model over-harvests the population
passed the tipping point consistently, resulting in the immediate crash
of the stock and thus derives minimal long-term catch.

The results across replicate stochastic simulations can most
easily be compared by using the relative differences in net present
value realized by each of the model. Figure 6 shows the distribution of values
realized by each simulation shown in Figure 5.  The GPDP
most consistently realizes a value close to the optimal solution,
and importantly avoids ever driving the system across the tipping point,
which results in the near-zero value cases in the parametric models.

@@

Changing the noise or the distance between stable and unstable points
does not impact the performance of the GP relative to the optimal solution
given the true model and true parameters.  The parametric models are
more senstive to this difference.  Large noise relative to the distance between the stable and unstable
point increases the chance of a stochastic transition.  More precisely,
if we let $L = K - x_c$, then the probability of a transition in a given
window of time $T$ scales as

@@

(following the Ahrenius relationship, see @Gardiner2009 for the derivation).
As a result, the impact of using a model which underestimates the risk of harvesting
past the critical point is much more severe, since this such a situation
occurs more often.  Conversely, with large enough distance between the
optimal escapement and unstable points relative to the noise, the chance
of a transition becomes vanishingly small and all models can be modeled
near-optimally. Models that underestimate the cost incurred by stock
sizes fluctuating significantly below the optimal escapement level will
not perform poorly as long as those fluctuations are sufficently rare.

Figure 7 shows that the GPDP is little effected by increasing noise or increasing
Allee effects over much of the range. Much larger stochasticity or higher Allee
effects make even the optimal solution without any model or parameter uncertainty
unable to harvest the population effectively (e.g. noise large enoug to violate the
self-sustaining criterion of @Reed1979). Only at the largest Allee effect
sizes are some replicates seen to deviate more from the optimal solution.

```{r figure_7, fig.width=6, dependson=c("sensitivity-trends"), fig.cap="The effect of increasing noise or decreasing Allee threshold levels on the net present value of the fishery when managed under the GPDP, relative to managing under the true model (with known parameters).  Other than the focal parameter (noise, Allee threshold), other parameters are held fixed as above to illustrate this effect."}

@@

>>>>

abstract: |
          Model uncertainty and limited data are fundamental
          challenges to robust management of human intervention in a 
          natural system.  These challenges
          are acutely highlighted by concerns that many ecological systems
          may contain tipping points, such as Allee population sizes.  
          Before a collapse, we do not know
          where the tipping points lie, if they exist at all.  Hence,
          we know neither a complete model of the system dynamics nor
          do we have access to data in some large region of state-space
          where such a tipping point might exist.  We illustrate
          how a Bayesian Non-Parametric (BNP) approach using a Gaussian Process (GP) prior
          provides a flexible representation of this inherent
          uncertainty. We embed GPs in a
          Stochastic Dynamic Programming (SDP) framework in order to make
          robust management predictions with both model uncertainty and
          limited data. We use simulations to evaluate this
          approach as compared with the standard approach of using model
          selection to choose from a set of candidate models.  We find that model selection erroneously
          favors models without tipping points -- leading to harvest policies
          that guarantee extinction. The GPDP 
          performs nearly as well as the true model and significantly
          outperforms standard appraoches. We illustrate this using examples
          of simulated single-species dynamics, where the standard model
          selection approach should be most effective, and find that it
          still fails to account for uncertainty appropriately and leads
          to population crashes, while management based on the GPDP 
          does not, since it does not underestimate the uncertainty
          outside of the observed data.

@@

Decision making under uncertainty is a ubiquitous challenge in the
management of human intervention in natural resources and conservation.
Ecological dynamics are frequently complex and difficult to measure,
making uncertainty in our understanding a persistent challenge.
Decision-theoretic approaches provide a framework to determine the
best sequence of actions in face of uncertainty, but only when that
uncertainty can be meaningfully quantified [@Fischer2009].  Over the last
four decades (but stretching even further back in time), marked by  the
seminal contributions of @Clark1976, @Clark2009 and @Walters1978,
dynamic optimization methods, particularly Stochastic Dynamic Programming
(SDP), have become increasingly important as a means of
understanding how to management human intervention into natural systems.
Concomittantly, there has been increasing recognition of the importance of
Allee population sizes or `tipping points' [@Scheffer2001;
@Polasky2011] that management actions may inadvertently
cause populations to cross.  @Scheffer2009 emphasized the difficulties of
formulating even qualitatively correct models of the underlying processes.

We develop a novel approach to address these concerns in the context of
fisheries; though the underlying challenges and methods are germane to
many other problems of conservation or natural resource exploitation,
such as forestry.  Economic value and ecological concern have made
marine fisheries the crucible for much of the founding work [@Gordon1954;
@Clark1976; -@Clark2009; @Reed1979; @May1979; @Ludwig1982] on management
under uncertainty.  Global declines [@Worm2006] and the controversy
surrounding their interpretation [@Hilborn2007; @Worm2009] make
understanding these challenges all the more pressing.

@@

The dual concerns of model uncertainty and
incomplete data coverage pose a substantial challenge to existing decision-theoretic
approaches [@Brozovic2011].  Because intervention is often (but not always, see @Hughes2013) too late after a tipping point has been crossed, management
is most often concerned with avoiding potentially catastrophic tipping
points before any data are available at or following a transition that
would more clearly reveal these regime shift dynamics [e.g. @Bestelmeyer2012].

We illustrate how a Stochastic Dynamic Programming (SDP) algorithm
[@Mangel1988; @Marescot2013] can be driven by the
predictions from a Bayesian Non-Parametric (BNP) model of population dynamics [@Munch2005a].
This provides two distinct advantages compared with contemporary
approaches.  First, using a BNP sidesteps the need for an accurate
model-based description of the system dynamics.  Second, the BNP can
better reflect uncertainty that arises when extrapolating a model outside
of the data on which it was fit.  We illustrate that when the correct
model is not known, this latter feature is crucial to providing a robust
decision-theoretic approach in face of substantial structural uncertainty.

This paper represents the first ecological application of the SDP
decision-making framework without an a priori model of the underlying
dynamics.  In contrast to parametric models which can only reflect
uncertainty in parameter estimates, the BNP approach provides a 
state-space dependent representation of uncertainty.  This permits
a much greater uncertainty far from the observed data than near the
observed data.  These features allow the Gaussian Process Dynamic Programming (GPDP) approach to find robust
management solutions in face of limited data and without knowledge of
the correct model structure.

@@

We let $X_t$ denote the size (numbers or biomass) of the focal population at time $t$ and assume that in the absence of take its dynamics are. NOTE: EQUATION NUMBERS ARE ALL MESSED UP. THIS SHOULD BE EQ 1, BUT DOES NOT APPEAR ON THE PDF YOU SENT TO ME

$$ X_{t+1}= Z_t f(X_t) $$

@@

In order to find that policy, we introduce the value function$V_t(x_t)$ representing the total discounted catch from time $t$ onwards given that $X_t=x$.  This value function satisfies an equation of SDP  [ @Mangel1988; @Clark2000; @Clark2009], sometimes called the Bellman-Hamilton-Jacobi equation [@Mangel2014],

@@

where expectation is taken over all possible values of the next state, $X_t{+1}$, and maximized over all possible choices of harvest, $h_t$.  That is, at time $t$, when population size is $x$ and harvest $h_t$ is applied, the immediate return is $R(h_t, X_t)$. In the classical case, the sole source of uncertainty is the process stochasticity term, $Z$, and thus the expectation above is equivalent to taking expectations over $Z$. That is

@@

We consider the finite time problem with $T=$ `r MaxT`, which we solve using value iteration algorithm [@Mangel1988; @Clark2000].

@@

The core difference for our purpose between the GP representation and the parametric
models is that the GP model is defined explicitly in reference to 
some observed data.  Unlike the models above, it cannot be specified by
the value of some parameters alone.  Although called nonparametric, the
GP (a)  still involves the estimation of
parameters, and (b), statisticians also use non-parametric to mean models
that make only weak distributional assumptions  [e.g. @Lehmann1975].
However, we retain the terminology for consistency with previous studies.

@@

We assume that the data $X_o$ are observed with some process stochasticity,

@@

where $\varepsilon$ are IID normal random variables with zero-mean and variance $\sigma_n^2$. Note that we have chosen to
assume additive stochasticity.  While we could just as easily consider log-normal stochasticity
as in the parametric models, we make this choice to bring home the point that the
Gaussian process approach need not have structurally correct noise form either.

@@

We use vague inverse Gamma priors on both the lengthscale $\ell$ and $\sigma$, of the form

@@

Using the simulated training data we also estimate a Gaussian process
defined by a radial basis function kernel of two parameters: $\ell$,
which gives the characteristic length-scale over which correlation between
two points in state-space decays, and $\sigma$, which gives the scale
of the process noise by which observations $Y_{t+1}$ may differ from
their predicted values $X_{t+1}$ given an observation of the previous
state, $X_t$. @Munch2005a give an accessible introduction to the use
of Gaussian processes in providing a Bayesian nonparametric description
of the stock-recruitment relationship.

@@

We use a Metropolis-Hastings Markov Chain Monte Carlo (@Gelman2003) to
infer posterior distributions of the two parameters of the GP (Figure
S13, code in appendix), under weakly informative Gaussian priors
(see parameters in table S5). Since the posterior distributions differ
substantially from the priors (Figure S13), we can be assured that most
of the information in the posterior comes from the data rather than the
prior belief.

@@

The method of Gaussian Process Dynamic Programming (GPDP)
--------------------------------------------------

@@

To ensure our results are robust to our choice of parameters, 
consider 96 different scenarios, described in detail below.   To begin, to help  better
understand the process, we first describe in detail the
results of a single scenario.

@@

All models fit the observed data rather closely and with relatively small uncertainty, as illustrated in the posterior predictive curves in Figure 1, in which we show the training data of stock sizes observed over time as points, overlaid with the step-ahead predictions of each estimated model using the parameters sampled from their posterior distributions.  Each model manages to fit the observed data rather closely. Compared to the true model most estimates appear to over-fit, predicting fluctuations that are actually due purely to stochasticity.  Model selection criteria (Table 1) penalize more complex models and show a preference for the simpler Ricker model over the  models with alternative stable states (Allen and Myers).  Details on MCMC estimates for each model, traces, and posterior distributions can be found in the appendix.

NOTE: IN THE PDF YOU SENT THERE IS WEIRD STUFF ON LINE 237

@@

```{r figure_2, dependson=c("assemble-models", "plot-options"), fig.cap="The inferred Gaussian process compared to the true process and maximum-likelihood estimated process.  We show the expected value for the function $f$ under each model.  Two standard deviations from the estimated Gaussian process covariance with (light grey) and without (darker grey) measurement error are also shown.  The training data are also shown as black points.  The GP is conditioned on (0,0), shown as a pseudo-data point.", fig.width=8, fig.height=6}
x_grid_short <- x_grid[1:40]
gp_short <- gp_predict(gp, x_grid_short, burnin=1e4, thin=300)
models_posteriors <- 
  melt(lapply(sample(100, 50), 
              function(i){
    sample_gp <- mvrnorm(1, 
                            gp_short$Ef_posterior[,i],         
                            gp_short$Cf_posterior[[i]])
    data.frame(stock = x_grid_short, 
               GP = sample_gp,
               y = sample_gp,
               ymin = sample_gp - 2 * sqrt(gp_short$E_Vf), 
               ymax = sample_gp + 2 * sqrt(gp_short$E_Vf), 
               True = sapply(x_grid_short,f,0, p),  
               MLE = sapply(x_grid_short,f,0, est$p), 
               Allen = sapply(x_grid_short, allen_f, 0, pardist[i,]), 
               Ricker = sapply(x_grid_short, ricker_f, 0, ricker_pardist[i,]), 
               Myers = sapply(x_grid_short, myers_f, 0, myers_pardist[i,]))
             }), 
       id=c("stock", "y", "ymin", "ymax"))
ggplot(models_posteriors) + 
    geom_ribbon(aes(x=stock, y=y, ymin=ymin, ymax=ymax, group=L1), 
                  fill = "gray80", 
                  data=subset(models_posteriors, variable == "GP")) + 
    geom_line(aes(stock, value, col = variable, 
                  group=interaction(L1,variable)), 
              alpha=.2) + 
    geom_point(data = obs, aes(x,y), alpha = 0.8) + 
    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
    facet_wrap(~variable) + 
    scale_colour_manual(values=colorkey) +  
    theme(legend.position="none")

@@

We show the  mean inferred state space dynamics of each model relative to the
true model used to generate the data in Figure 2, predicting
the relationship between observed stock size (x-axis) to the stock
size after recruitment the following year.  Note that in contrast to
the  other models shown, the expected Gaussian process corresponds to a
distribution of curves - as indicated by the gray band - which itself
has a mean shown in black. Parameter uncertainty (not shown) spreads
out the estimates further.  The observed data from which each model are
estimated is also shown.  The observations come from only a limited region
of state space corresponding to unharvested or weakly harvested system.
No observations occur at the theoretical optimum harvest rate or near
the tipping point.

NOTE: THERE IS NO MENTION OF FIGURE 3 IN THE TEXT

@@

Despite the similarities in model fits to the observed data, the
policies inferred under each model differ widely (Figure 4).
Policies are shown in terms of target escapement, $S_t$.  Under models
such as this a constant escapement policy is expected to be optimal
[@Reed1979], whereby population levels below a certain size $S$ are
unharvested, while above that size the harvest strategy aims to return the
population to $S$, resulting in the hockey-stick shaped policies shown.
Only the structurally correct model (Allen model) and the GP produce
policies close to the true optimum policy (where both the underlying
model structure and parameter values are known without error).

```{r figure_5, dependson=c("sim"), fig.cap="In the management context, GPDP outperforms approaches based on parametric models. We show 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6}
ggplot(sims_data) + 
  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) + 
  facet_wrap(~method) + guides(legend.position="none")

@@

In Figure 5, we show the consequences of managing 100 replicate realizations of the simulated
fishery under each of the policies estimated..
As anticipated from the policy curves, the structurally correct model
under-harvests, leaving the stock to vary around its unfished optimum.
The structurally incorrect Ricker model over-harvests the population
passed the tipping point consistently, resulting in the immediate crash
of the stock and thus derives minimal long-term catch.

The results across replicate stochastic simulations can most
easily be compared by using the relative differences in net present
value realized by each of the model (Figure 6).  The GPDP
most consistently realizes a value close to the optimal solution,
and importantly avoids ever driving the system across the tipping point,
which results in the near-zero value cases in the parametric models.

@@

Changing the intensity of the stochasticity or the distance between stable and unstable points
does not impact the performance of the GP relative to the optimal solution
given the true model and true parameters.  The parametric models are
more senstive to this difference.  Large noise relative to the distance between the stable and unstable
point increases the chance of a stochastic transition.  More precisely,
if we let $L = K - x_c$, then the probability of a transition in a given
window of time $T$ scales as

@@

(see @Gardiner2009 or @Mangel2006 for the derivation).
As a result, the impact of using a model which underestimates the risk of harvesting
past the critical point is much more severe, since this such a situation
occurs more often.  Conversely, with large enough distance between the
optimal escapement and unstable points relative to the noise, the chance
of a transition becomes vanishingly small and all models can be modeled
near-optimally. Models that underestimate the cost incurred by stock
sizes fluctuating significantly below the optimal escapement level will
not perform poorly as long as those fluctuations are sufficently rare.

The GPDP is weakly influenced by increasing stochasticity or increasing
Allee effects over much of the range (Figure 7) Much larger stochasticity or higher Allee
effects make even the optimal solution without any model or parameter uncertainty
unable to harvest the population effectively (e.g. noise large enoug to violate the
self-sustaining criterion of @Reed1979). Only at the largest Allee effect
sizes are some replicates seen to deviate more from the optimal solution.

```{r figure_7, fig.width=6, dependson=c("sensitivity-trends"), fig.cap="The effect of increasing noise or decreasing Allee threshold levels on the net present value of the fishery when managed under the GPDP, relative to managing under the true model (with known parameters).  Other than the focal parameter (stochasticity, Allee threshold), other parameters are held fixed as above to illustrate this effect."}

@@

NOTE: THE PDF HAS TWO }} IN LINE 351
